---
title: "Recommended Workflow for using devMSMs with Longitudinal Data"
author: "`Isabella C. Stallworthy, Meriah L. DeJoseph, Emily R. Padrutt, Noah Greifer, Daniel Berry`")
date: "`r Sys.Date()`"
output: html_document
---

This example workflow provides more detail on the application of *devMSMs* to the empirical example in the manuscript, xxxx, which can be found here: xx.

Please review the accompanying manuscript for a full conceptual and practical introduction to MSMs in the context of developmental data. Please also see the vignettes on the *devMSMs* website for step-by-step guidance on the use of this code: <https://istallworthy.github.io/devMSMs/index.html>.

We highly recommend first implementing the *Data Requirements* and *Specify Core Inputs* Vignettes prior to using this workflow.

Headings denote accompanying website vignettes and steps that correspond to the manuscript Section 7. We suggest using the interactive outline tool (located above the Console) for ease of navigation.

The code in each code chunk is set up to showcase all possible inputs to each function (both required and optional) to aid the user's use of the full range of package functionality. Example possible values for the optional input are shown for each function.

Please see the website vignettes and/or type `?functionName` into the console for more guidance on the arguments for each function.

Functions from devMSMs: <https://github.com/istallworthy/devMSMs>

<br>

# *Installation*

<https://istallworthy.github.io/devMSMs/index.html>

Until *devMSMs* is available on CRAN, you will need to install it directly from Github (<https://github.com/istallworthy/devMSMs>), as shown below.

```{r}
install.packages("devtools", quiet = TRUE)
library(devtools)

install_github("istallworthy/devMSMs", quiet = TRUE)
library(devMSMs)

install_github("istallworthy/devMSMsHelpers", quiet = TRUE)
library(devMSMsHelpers)
```

<br>

# *Specify Core Inputs* and *Data Requirements & Inspection* Vignettes

## Phase 0: Preliminary Conceptual & Data Preparation

Please see the accompanying manuscript for steps P1 (creating hypotheses) and P2 (creating a DAG).

### STEP P3. Specify Core Inputs

The first step is to specify core inputs for use with the package.

In addition to the core inputs supplied directly to `initMSM()` in Step P4.4 (See the Specify Core Inputs vignette for more information including guidance on specifying core inputs P3.1-3 & P3.5), we specify the following additional inputs:

#### P3.4. Required for Continuous Exposures: Identify High and Low Cutoff Values

See the Specify Core Inputs vignette for more information.

Below, we specify the 60th and 30th percentiles to demarcate high and low levels of economic strain exposure, respectively.

```{r}
hi_lo_cut <- c(0.6, 0.3)
```

### P3.5 Optional: Exposure Epochs

We specify that the first two exposure time points (6 and 15 months) will be considered infancy, the second two (34 and 25 months) toddlerhood, and the final (58 months) childhood.

```{r}
epochs <- c("Infancy", "Infancy", "Toddlerhood", "Toddlerhood", "Childhood")
```

#### P3.6. Recommended: Specify Hypotheses-Relevant Exposure Histories

See the Specify Core Inputs vignette for more information.

Below, we specify all reference histories and all comparison histories associated with our two hypotheses. We will specify more specific reference and comparison histories corresponding to each hypothesis in Step 6.

```{r}
reference <- c("h-l-l", "l-h-h", "l-l-l", "l-h-l", "l-l-h", "h-h-h")

comparison <- c("l-l-h", "l-h-l", "l-l-l", "h-l-h", "h-h-l", "h-l-l", "l-h-h")
```

#### P3.7 Required: Specify Outcome

```{r}
outcome <-  "StrDif_Tot.58"
```

<br>

### STEP P4. Data Preparation & Inspection

Please see the ExampleDataPrep.rmd file for more details on data preparation for this empirical example. We also highly recommend first implementing the *Data Preparation Vignette* <https://istallworthy.github.io/devMSMs/articles/Preliminary_Steps.html> for preparing data starting in different formats (P4.1 - P4.3a).

#### P4.3b. Required: Read in Wide Data

Please see the *Data Requirements* vignette for guidance on preparing complete data in wide format

*devMSMs* requires data in one of the following wide data formats (see Figure 1) for use in the package:

-   a single data frame of data in wide format with no missing data

-   a mids object (output from mice::mice()) of data imputed in wide format

-   a list of data imputed in wide format as data frames\
    <br>

We first read in data that has been formatted and multiply imputed 5 times using the *mice* package using the random forest imputation technique.

```{r}
data <- readRDS("/Users/isabella/Library/CloudStorage/Box-Box/BSL General/MSMs/testing/isa/imputations/ESETA1-StrDif_Tot.58_all_imp.rds") 
```

<br>

#### P4.4 Required: Create MSM Object

Below, we create an MSM object by supplying our imputed data, exposure, time invariant and time-varying confounders, exposure epochs, the separator preceding the time point for time-varying variables, and an optional home directory for saving output.

```{r}
set.seed(1234)

obj <- initMSM(
  data,
  exposure = c("ESETA1.6", "ESETA1.15", "ESETA1.24", "ESETA1.35", "ESETA1.58"), 
  ti_conf =  c("state", "BioDadInHH2", "PmAge2", "PmBlac2", "TcBlac2", "PmMrSt2", 
               "PmEd2", "KFASTScr", "RMomAgeU", "RHealth", "HomeOwnd", "SWghtLB", 
               "SurpPreg", "SmokTotl", "DrnkFreq",
               "peri_health", "caregiv_health", "gov_assist"),
  tv_conf = c("SAAmylase.6","SAAmylase.15", "SAAmylase.24", 
              "MDI.6", "MDI.15",                                            
              "RHasSO.6", "RHasSO.15", "RHasSO.24","RHasSO.35",                                         
              "WndNbrhood.6","WndNbrhood.24", "WndNbrhood.35",                                      
              "IBRAttn.6", "IBRAttn.15", "IBRAttn.24",                                   
              "B18Raw.6", "B18Raw.15", "B18Raw.24",                                           
              "HOMEETA1.6", "HOMEETA1.15", "HOMEETA1.24", "HOMEETA1.35",                              
              "InRatioCor.6", "InRatioCor.15", "InRatioCor.24", "InRatioCor.35",                         
              "CORTB.6", "CORTB.15", "CORTB.24",                                                                  
              "EARS_TJo.24", "EARS_TJo.35",                                        
              "LESMnPos.24", "LESMnPos.35",                                  
              "LESMnNeg.24", "LESMnNeg.35",       
              "StrDif_Tot.35", 
              "fscore.35"
              ),
  epoch = epochs,
  sep = "\\.",
  home_dir = '/Users/isabella/Library/CloudStorage/Box-Box/BSL General/MSMs/testing/isa' 
)
```

View the msm object

```{r}
print(obj)
```

#### P4.5 Recommended: Inspect Inspect Exposure Histories and Data

The helper `inspectData()` function outputs the following files into the home directory: a correlation plot of all variables in the dataset, tables of exposure and outcome descriptive statistics, and two summary tables of the confounders considered at each time point.

```{r, eval = FALSE}
inspectData(data = data, 
            obj = obj, 
            outcome = outcome, 
            hi_lo_cut = hi_lo_cut,
            reference = reference, 
            comparison = comparison, 
            verbose = TRUE,
            save.out = TRUE)
```

Here, we see summaries of the data types as well as reasonable cell counts in each of our specified histories, for each imputed dataset. <br> <br>

# *Workflow: Continuous Exposure* Vignette

<https://istallworthy.github.io/devMSMs/articles/Workflow_Continuous_Exposure.html>

## PHASE 1: Confounder Adjustment

The first phase of the MSM process is focused on eliminating confounding of the relation between exposure and outcome.

### STEP 1: Create Full Weights Formulas & Conduct Pre-Balance Checking

The first step is to create full weights formulas that reflect all measured confounders at each exposure time point.

#### 1a. Create Full Weights Formulas at each Exposure Time Point

Users have the option to specify concurrent confounders to retain and we recommend doing so consistently throughout this workflow.

Please see the *Customize Weights Formulas* Vignette at the link below for more detail on custom formulas. <https://istallworthy.github.io/devMSMs/articles/Customize_Balancing_Formulas.html>

```{r}
type <- "full"

full_formulas <- createFormulas(obj = obj, 
                                type = type, 
                                save.out = TRUE)
```

Inspect

```{r}
print(full_formulas)
```

#### 1b. Conduct Exploratory Pre-Balance Assessment

The next step is to examine initial imbalance between confounders and exposure prior to IPTW weighting. Users have the option to specify balance threshold(s), which we recommend doing consistently throughout this workflow.

We first specify two balance thresholds and a list of important confounders, which will be held to the higher (first) balancing threshold. These specifications will be used for all balance assessment.

```{r}
balance_thresh <- c(0.05, 0.1) 

imp_conf <- c("InRatioCor.6", "InRatioCor.15", "InRatioCor.24", "InRatioCor.35",
              "PmEd2") 
```

```{r}
prebalance_stats <- assessBalance(obj = obj, 
                                  data = data,
                                  balance_thresh = balance_thresh, 
                                  imp_conf = imp_conf, 
                                  save.out = TRUE)
```

Inspect and save out

```{r}
# summarize prebalance stats averaging across imputed datasets 
summary(prebalance_stats, 
        save.out = TRUE)

# view and save out balance stats averaged across imputed datasets
print(prebalance_stats,
      save.out = TRUE)

# save out plots for all time points and imputed datasets 
for (i in 1:5){
  for (t in 1:5){
    plot(prebalance_stats, 
         i = i, 
         t = t, 
         save.out = TRUE)
  }
}

# save out balance stats for all imputed datasets
for (i in 1:5){
    print(prebalance_stats, 
          i = i, 
          save.out = TRUE) 
}
```

<br>

### STEP 2: Create Simplified Weights Formulas & Determine Optimal Weighting Method

The next step is to specify shorter, simplified weights formula for the purposes of determining the weighting method optimal for the data.

#### 2a. Create Simplified Wights Formulas

First, create shorter, simplified weights formulas at each exposure time point.

```{r}
type <- "short" 

short_formulas <- createFormulas(obj = obj, 
                                 type = type, 
                                 save.out = TRUE) 
```

Inspect

```{r}
print(short_formulas)
```

#### 2b. Create IPTW Balancing Weights Using Multiple Weighting Methods

We recommend users use the short formulas to create IPTW weights using all the available weighting methods: "glm", "gbm", "bart", "super", and "cbps."

```{r}
formulas <- short_formulas

method <- "cbps" 

weights.cbps <- createWeights(obj = obj, 
                              data = data, 
                              formulas = formulas, 
                              method = method, 
                              # maxit = 1, # for testing purposes only; makes it run faster 
                              save.out = TRUE)
# Read in if you've run them already:
# weights.cbps <- readRDS("/Users/isabella/Library/CloudStorage/Box-Box/BSL General/MSMs/testing/isa/weights/type_short-exposure_ESETA1-method_cbps.rds")
```

Inspect by imputed dataset

```{r}
print(weights.cbps, 
      i = 1)

plot(weights.cbps, 
     i = 1, 
     save.out = TRUE)

summary(weights.cbps[[1]]) 
```

We then fit weights for all other available weighting methods.

```{r}
method <- "glm"
weights.glm <- createWeights(obj = obj, 
                             data = data, 
                             formulas = formulas, 
                             method = method,
                             save.out = TRUE)
print(weights.glm, 
      i = 2)
plot(weights.glm, 
     i = 1,
     save.out = TRUE)
```

```{r}
method <- "gbm"
weights.gbm <- createWeights(obj = obj, 
                             data = data, 
                             formulas = formulas, 
                             method = method,
                             save.out = TRUE)
# Read in if you've run already:
# weights.gbm <- readRDS("/Users/isabella/Library/CloudStorage/Box-Box/BSL General/MSMs/testing/isa/weights/type_short-exposure_ESETA1-method_gbm.rds")

print(weights.gbm, 
      i = 2)
plot(weights.gbm, 
     i = 1, 
     save.out = TRUE)
summary(weights.gbm[[1]]) 
```

```{r}
method <- "bart"
weights.bart <- createWeights(obj = obj, 
                              data = data, 
                              formulas = formulas, 
                              method = method,
                              save.out = TRUE)
print(weights.bart, 
      i = 2)
plot(weights.bart, 
     i = 1, 
     save.out = TRUE)
```

```{r}
method <- "super"
weights.super <- createWeights(obj = obj, 
                               data = data, 
                               formulas = formulas, 
                               method = method,
                               save.out = TRUE)
print(weights.super, 
      i = 2)
plot(weights.super, 
     i = 1, 
     save.out = TRUE)
```

#### 2c. Assess All Weighting Methods to Determine Optimal Method

The next step is to assess balance for each weighting method and for the user to determine the optimal weighting method.

```{r}
weights <- weights.cbps 

balance_stats.cbps <- assessBalance(data = data, 
                                    obj = obj, 
                                    weights = weights,
                                    imp_conf = imp_conf, 
                                    balance_thresh = balance_thresh,
                                    save.out = TRUE)

# to load in from file (output from saving above!)
# balance_stats.cbps <- readRDS("/Users/isabella/Library/CloudStorage/Box-Box/BSL General/MSMs/testing/isa/balance/weighted-exposure_A-method_cbps.rds")
```

```{r}
weights <- weights.glm
balance_stats.glm <- assessBalance(data = data, 
                                   obj = obj, 
                                   weights = weights,
                                   imp_conf = imp_conf, 
                                   balance_thresh = balance_thresh,
                                   save.out = TRUE)
```

```{r}
weights <- weights.gbm
balance_stats.gbm <- assessBalance(data = data, 
                                   obj = obj, 
                                   weights = weights,
                                   imp_conf = imp_conf, 
                                   balance_thresh = balance_thresh,
                                   save.out = TRUE)
```

```{r}
weights <- weights.bart
balance_stats.bart <- assessBalance(data = data, 
                                    obj = obj, 
                                    weights = weights,
                                    imp_conf = imp_conf, 
                                    balance_thresh = balance_thresh,
                                    save.out = TRUE)
```

```{r}
weights <- weights.super 
balance_stats.super <- assessBalance(data = data, 
                                     obj = obj, 
                                     weights = weights,
                                     imp_conf = imp_conf, 
                                     balance_thresh = balance_thresh,
                                     save.out = TRUE)
```

We then compare the balance of the different weighting methods.

```{r}
summary(balance_stats.cbps, 
        save.out = TRUE)
print(balance_stats.cbps, 
      save.out = TRUE)

summary(balance_stats.glm, 
        save.out = TRUE)
print(balance_stats.glm, 
      save.out = TRUE)

summary(balance_stats.gbm, 
        save.out = TRUE)
print(balance_stats.gbm, 
      save.out = TRUE)

summary(balance_stats.bart, 
        save.out = TRUE)
print(balance_stats.bart, 
      save.out = TRUE)

summary(balance_stats.super, 
        save.out = TRUE)
print(balance_stats.super, 
      save.out = TRUE)
```

We identify CBPS as the weighting method that most successfully reduces imbalance.

<br>

### STEP 3: Create Updated Formulas & Re-Specify Weights Using Optimal Weighting Method

The next step is to update the short weights formulas with any imbalanced confounders, and re-specify the IPTW weights.

#### 3a. Examine Balance of Optimal Weighting Method

First, assess how well each of the IPTW achieve balance for all measured confounders.

```{r}
# save all time points averaged across imps
for (t in 1:5){
  plot(balance_stats.glm, 
       t = t, 
       save.out = TRUE)
}

# save for all imps 
for (t in 1:5){
  for (i in 1:5){
  plot(balance_stats.glm, 
       t = t, 
       i = i, 
       save.out = TRUE)
  }
}


# save all time points averaged across imps
for (t in 1:5){
  plot(balance_stats.gbm, 
       t = t, 
       save.out = TRUE)
}

# save for all imps 
for (t in 1:5){
  for (i in 1:5){
  plot(balance_stats.gbm, 
       t = t, 
       i = i, 
       save.out = TRUE)
  }
}


# save all time points averaged across imps
for (t in 1:5){
  plot(balance_stats.super, 
       t = t, 
       save.out = TRUE)
}

# save for all imps 
for (t in 1:5){
  for (i in 1:5){
  plot(balance_stats.super, 
       t = t, 
       i = i, 
       save.out = TRUE)
  }
}

# save all time points averaged across imps
for (t in 1:5){
  plot(balance_stats.bart, 
       t = t, 
       save.out = TRUE)
}

# save for all imps 
for (t in 1:5){
  for (i in 1:5){
  plot(balance_stats.bart, 
       t = t, 
       i = i, 
       save.out = TRUE)
  }
}




summary(balance_stats.cbps)

plot(balance_stats.cbps, 
     t = 1, 
     save.out = FALSE)

# save all time points averaged across imps
for (t in 1:5){
  plot(balance_stats.cbps, 
       t = t, 
       save.out = TRUE)
}

# save for all imps 
for (t in 1:5){
  for (i in 1:5){
  plot(balance_stats.cbps, 
       t = t, 
       i = i, 
       save.out = TRUE)
  }
}

print(balance_stats.cbps) 
```

#### 3b. Update Simplified Formulas

Next, update the short formulas with any imbalanced confounders.

```{r}
type <- "update"

bal_stats <- balance_stats.cbps

updated_formulas <- createFormulas(obj = obj, 
                                   type = type,
                                   bal_stats = bal_stats,
                                   save.out = TRUE)
```

Inspect

```{r}
print(updated_formulas)
```

#### 3c. Create Final Balancing Weights

Next, create final balancing weights using the optimal weighting method and updated weights formulas.

```{r}
formulas <- updated_formulas

method <- "cbps" 

final_weights <- createWeights(data = data, 
                               obj = obj, 
                               method = method, 
                               formulas = formulas,
                               # maxit = 1, # just for testing purposes 
                               save.out = TRUE)
# final_weights <- readRDS("/Users/isabella/Library/CloudStorage/Box-Box/BSL General/MSMs/testing/isa/weights/type_update-exposure_ESETA1-method_cbps.rds")
```

Inspect

```{r}
print(final_weights, 
      i = 1)

summary(final_weights[[1]])

plot(final_weights, 
     i = 1,
     save.out = TRUE)
```

#### 3d. Trim Final Balancing Weights

Next, trim the final balancing weights to reduce the heavy right tail.

##### Main

First, trim the main weights.

```{r}
quantile <- 0.95 

weights <- final_weights
trim_weights <- trimWeights(weights = weights, 
                            at = quantile, 
                            save.out = TRUE)
```

Inspect

```{r}
print(trim_weights, 
      i = 1)

plot(trim_weights, 
     i = 1, 
     save.out = TRUE)
```

##### Sensitvity Analyses

Next, conduct sensitivity analyses using two other quantile values.

```{r}
quantile <- 0.92 

trim_weights.s1 <- trimWeights(weights = weights, 
                               at = quantile,
                               save.out = TRUE)
plot(trim_weights.s1, 
     i = 1, 
     save.out = TRUE)
```

```{r}
quantile <- 0.98

trim_weights.s2 <- trimWeights(weights = weights, 
                               at = quantile,
                               save.out = TRUE)
plot(trim_weights.s2, 
     i = 1, 
     save.out = TRUE)
```

<br>

### STEP 4: Conduct Final Balance Assessment

Next, conduct a final balance assessment using all measured confounders (i.e., the full weights formulas).

#### Trimmed

Assess balance of trimmed weights

```{r}
weights <- trim_weights

final_balance_stats_trim <- assessBalance(data = data, 
                                     obj = obj, 
                                     balance_thresh = balance_thresh,
                                     imp_conf = imp_conf,
                                     weights = weights,
                                     save.out = TRUE)
```

Summarize and inspect

```{r}
summary(final_balance_stats_trim, 
        save.out = TRUE)

summary(trim_weights[[1]])


# average ESS
mean(summary(trim_weights[[1]])$ESETA1.6$effective.sample.size$Total[2],
     summary(trim_weights[[2]])$ESETA1.6$effective.sample.size$Total[2],
     summary(trim_weights[[3]])$ESETA1.6$effective.sample.size$Total[2],
     summary(trim_weights[[4]])$ESETA1.6$effective.sample.size$Total[2],
     summary(trim_weights[[5]])$ESETA1.6$effective.sample.size$Total[2]
)


# plot(final_balance_stats_trim, 
#      t = 2, 
#      save.out = TRUE)
# 
# print(final_balance_stats_trim, 
#       save.out = TRUE)
```

#### Untrimmed

Assess balance of untrimmed weights

```{r}
weights <- final_weights

final_balance_stats_untrim <- assessBalance(data = data, 
                                     obj = obj, 
                                     balance_thresh = balance_thresh,
                                     imp_conf = imp_conf,
                                     weights = weights,
                                     save.out = TRUE)
```

Summarize and inspect

```{r}
summary(final_balance_stats_untrim, 
        save.out = TRUE)
summary(final_weights[[1]])

# plot(final_balance_stats_untrim, 
#      t = 2, 
#      save.out = TRUE)
# 
# print(final_balance_stats_untrim, 
#       save.out = TRUE)
```

#### Sensitvity Analyses

Next, conduct the recommended specifying sensitivity analyses to match the main analyses above.

```{r}
weights <- trim_weights.s1

final_balance_stats.s1 <- assessBalance(data = data, 
                                        obj = obj,
                                        weights = weights,
                                        balance_thresh = balance_thresh, 
                                        imp_conf = imp_conf,
                                        save.out = "balance-weighted-exposure_ESETA1-method_cbps-form_update-s1.rds")
summary(final_balance_stats.s1, 
        save.out = TRUE)
summary(trim_weights.s1[[1]])
```

```{r}
weights <- trim_weights.s2

final_balance_stats.s2 <- assessBalance(data = data, 
                                        obj = obj, 
                                        weights = weights,
                                        balance_thresh = balance_thresh, 
                                        imp_conf = imp_conf,
                                        save.out = "balance-weighted-exposure_ESETA1-method_cbp-form_update-s2.rds")
summary(final_balance_stats.s2, 
        save.out = TRUE)
summary(trim_weights.s2[[1]])

```

<br> <br>

## PHASE 2: Assess Substantive Associations between Exposure and Outcome

Lastly, having attenuated confounder associations, we model substantive associations.

### STEP 5: Fit Weighted Outcome Model 
First, select and fit a marginal outcome model.

#### 5a. Select a Weighted Outcome Model

##### Main

Set parameters

```{r}
family <- gaussian

link <- "identity" 

weights <- trim_weights

m <- "m0"

outcome <- "StrDif_Tot.58"
```

### 5b. Fit the Weighted Outcome Model

Fit weighted model

```{r}
models <- fitModel(data = data, 
                   obj = obj, 
                   weights = weights, 
                   outcome = outcome, 
                   model = m,
                   save.out = TRUE) 
```

(Fit unweighted model for illustration purposes only)

```{r}
models_unweighted <- fitModel(data = data,
                   obj = obj,
                   weights = NULL,
                   outcome = outcome,
                   model = m,
                   save.out = TRUE)
```

Inspect

```{r}
print(models, 
      save.out = TRUE)

print(models, 
      i = 3,
      save.out = TRUE)


# # unweighted
# print(models_unweighted, 
#       save.out = "fit_model_summary-outcome_StrDif_Tot_58-exposure_ESETA1-model_m0-all_avg_UNWEIGHTED.docx")
```

##### Sensitvity Analyses

Next, fit the recommended specifying sensitivity analyses to match the main analyses above.

```{r}
weights <- trim_weights.s1

models.s1 <- fitModel(data = data, 
                      obj = obj, 
                      weights = weights, 
                      outcome = outcome, 
                      model = m,
                      save.out = sprintf("outcome_StrDif_Tot-exposure_ESETA1-model_%s_s1.rds", m))
print(models.s1, 
      save.out = sprintf("fit_model_summary-outcome_StrDif_Tot_58-exposure_ESETA1-model_%s-all_avg_s1.docx", m))
```

```{r}
weights <- trim_weights.s2

models.s2 <- fitModel(data = data, 
                      obj = obj, 
                      weights = weights, 
                      outcome = outcome, 
                      model = m,
                      save.out = sprintf("outcome_StrDif_Tot-exposure_ESETA1-model_%s_s2.rds", m))

print(models.s2,
      save.out = sprintf("fit_model_summary-outcome_StrDif_Tot_58-exposure_ESETA1-model_%s-all_avg_s2.docx", m))
```

## Step 6. Estimate, Compare, and Visualize Model-Predicted Outcome as a Function of Exposure History

Lastly, estimate and compare user-specified exposure histories.

Set parameters

```{r}
hi_lo_cut <- c(0.6, 0.3) 

mc_comp_method <- "BH" 

dose_level <- "h"

exp_lab <- "Economic Strain" 

out_lab <- "Behavior Problems"

colors <- c("Dark2") 

model <- models 
```

### 6a & 6b Estimate, Compare, & Plot Potential Outcomes Across Exposure Histories

#### Timing hypotheses

Assess timing hypotheses about the role of exposure during infancy.

##### 1 dose timing

weighted

```{r}
reference_t1 <- "h-l-l" 
comparison_t1 <- c("l-l-h", "l-h-l", "l-l-l")

results_t1 <- compareHistories(fit = model, 
                            comparison = comparison_t1,
                            reference = reference_t1,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)
summary(results_t1)

print(results_t1, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-t1.docx")
plot(results_t1,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-t1.jpeg")
```

(Unweighted for illustrative purposes only)

```{r}
results_t1_uw <- compareHistories(fit = models_unweighted,
                            comparison = comparison_t1,
                            reference = reference_t1,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method,
                            dose = dose_level,
                            save.out = TRUE)
summary(results_t1_uw)

print(results_t1,
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-t1-uw.docx")
plot(results_t1,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-t1-uw.jpeg")
```

##### 2 dose timing

weighted

```{r}
reference_t2 <- "l-h-h" 
comparison_t2 <- c("h-l-h", "h-h-l") 

results_t2 <- compareHistories(fit = model, 
                            comparison = comparison_t2,
                            reference = reference_t2,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)
summary(results_t2)

print(results_t2, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-t2.docx")
plot(results_t2,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-t2.jpeg")
```

(Unweighted for illustrative purposes only)

```{r}
results_t2_uw <- compareHistories(fit = models_unweighted,
                            comparison = comparison_t2,
                            reference = reference_t2,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method,
                            dose = dose_level,
                            save.out = FALSE)
summary(results_t2_uw)

print(results_t2_uw,
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-t2-uw.docx")
plot(results_t2_uw,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-t2-uw.jpeg")
```

#### Dose hypotheses

Assess dose hypotheses.

##### 0 vs 1

weighted

```{r}
reference_d0 <- "l-l-l"

comparison_d0 <- c("h-l-l", "l-h-l", "l-l-h")

results_d0 <- compareHistories(fit = model, 
                            comparison = comparison_d0,
                            reference = reference_d0,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d0)

print(results_d0, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d0.docx")
plot(results_d0,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d0.jpeg")
```

(Unweighted for illustrative purposes only)

```{r}
results_d0_uw <- compareHistories(fit = models_unweighted,
                            comparison = comparison_d0,
                            reference = reference_d0,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method,
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d0_uw)

print(results_d0_uw,
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d0-uw.docx")
plot(results_d0_uw,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d0-uw.jpeg")
```

##### 1 vs 2

weighted

```{r}
reference_d1 <- c("h-l-l", "l-h-l", "l-l-h")

comparison_d1 <- c("h-h-l", "l-h-h", "h-l-h")

results_d1 <- compareHistories(fit = model, 
                            comparison = comparison_d1,
                            reference = reference_d1,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d1)

print(results_d1, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d1.docx")
plot(results_d1,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d1.jpeg")
```

(Unweighted for illustrative purposes only)

```{r}
results_d1_uw <- compareHistories(fit = models_unweighted,
                            comparison = comparison_d1,
                            reference = reference_d1,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method,
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d1_uw)

print(results_d1_uw,
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d1-uw.docx")
plot(results_d1_uw,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d1-uw.jpeg")
```

##### 2 vs 3

weighted

```{r}
reference_d2 <- "h-h-h"

comparison_d2 <- c("h-h-l", "l-h-h", "h-l-h")

results_d2 <- compareHistories(fit = model, 
                            comparison = comparison_d2,
                            reference = reference_d2,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d2)

print(results_d2, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d2.docx")
plot(results_d2,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d2.jpeg")
```

(Unweighted for illustrative purposes only)

```{r}
results_d2_uw <- compareHistories(fit = models_unweighted,
                            comparison = comparison_d2,
                            reference = reference_d2,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method,
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d2_uw)

print(results_d2_uw,
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d2-uw.docx")
plot(results_d2_uw,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d2-uw.jpeg")
```

##### 3 vs 1

weighted

```{r}
reference_d3 <- "h-h-h"

comparison_d3 <- c("h-l-l", "l-h-l", "l-l-h")

results_d3 <- compareHistories(fit = model, 
                            comparison = comparison_d3,
                            reference = reference_d3,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)

summary(results_d3)

print(results_d3, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d3.docx")
plot(results_d3,
     colors = c("#7570B3", "#D95F02", "#D95F02", "#D95F02"),
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d3.jpeg")
```

example sensitivity analyses

```{r}
results_d3.s1 <- compareHistories(fit = models.s1, 
                            comparison = comparison_d3,
                            reference = reference_d3,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d3.s1)

print(results_d3.s1, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d3.s1.docx")
plot(results_d3.s1,
     colors = c("#7570B3", "#D95F02", "#D95F02", "#D95F02"),
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d3.s1.jpeg")


results_d3.s2 <- compareHistories(fit = models.s2, 
                            comparison = comparison_d3,
                            reference = reference_d3,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d3.s2)

print(results_d3.s2, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d3.s2.docx")
plot(results_d3.s2,
     colors = c("#7570B3", "#D95F02", "#D95F02", "#D95F02"),
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d3.s2.jpeg")
```

(Unweighted for illustrative purposes only)

```{r}
results_d3_uw <- compareHistories(fit = models_unweighted,
                            comparison = comparison_d3,
                            reference = reference_d3,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method,
                            dose = dose_level,
                            save.out = FALSE)

summary(results_d3_uw)

print(results_d3_uw,
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d3-uw.docx")
plot(results_d3_uw,
     colors = c("#7570B3", "#D95F02", "#D95F02", "#D95F02"),
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d3-uw.jpeg")
```

##### 3 vs 0

weighted

```{r}
reference_d4 <- "h-h-h"

comparison_d4 <- "l-l-l"

results_d4 <- compareHistories(fit = model, 
                            comparison = comparison_d4,
                            reference = reference_d4,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d4)

print(results_d4, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d4.docx")
plot(results_d4,
     colors = c("#1B9E77", "#E7298A","#66A61E", "#E6AB02") ,#"#A6761D" "#666666"
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d4.jpeg")
```

Example sensitivity analyses

```{r}
results_d4.s1 <- compareHistories(fit = models.s1, 
                            comparison = comparison_d4,
                            reference = reference_d4,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d4.s1)

print(results_d4.s1, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d4.s1.docx")
plot(results_d4.s1,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d4.s1.jpeg")
```

```{r}
results_d4.s2 <- compareHistories(fit = models.s2, 
                            comparison = comparison_d4,
                            reference = reference_d4,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d4.s2)

print(results_d4.s2, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d4.s2.docx")
plot(results_d4.s2,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d4.s2.jpeg")
```

(Unweighted for illustrative purposes only)

```{r}
results_d4_uw <- compareHistories(fit = models_unweighted,
                            comparison = comparison_d4,
                            reference = reference_d4,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method,
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d4_uw)

print(results_d4_uw,
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d4-uw.docx")
plot(results_d4,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d4-uw.jpeg")
```

##### 2 vs 0

weighted

```{r}
reference_d5 <- "l-l-l"
comparison_d5 <- c("h-h-l", "l-h-h", "h-l-h")

results_d5 <- compareHistories(fit = model, 
                            comparison = comparison_d5,
                            reference = reference_d5,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method, 
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d5)

print(results_d5, 
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d5.docx")
plot(results_d5,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d5.jpeg")
```

(Unweighted for illustrative purposes only)

```{r}
results_d5_uw <- compareHistories(fit = models_unweighted,
                            comparison = comparison_d5,
                            reference = reference_d5,
                            hi_lo_cut = hi_lo_cut,
                            mc_comp_method = mc_comp_method,
                            dose = dose_level,
                            save.out = TRUE)
summary(results_d5_uw)

print(results_d5_uw,
      save.out = "comparisons_table-outcome_StrDif_Tot_58-exposure_ESETA1-d5-uw.docx")
plot(results_d5_uw,
     save.out = "comparisons_plot-outcome_StrDif_Tot_58-exposure_ESETA1-d5-uw.jpeg")
```

##### More stringent p-value correction considering all comparisons as one family

weighted

```{r}
all_results <- rbind(results_t1$comps,
                     results_t2$comps,
                     results_d0$comps,
                     results_d1$comps,
                     results_d2$comps,
                     results_d3$comps,
                     results_d4$comps,
                     results_d5$comps)
all_results$p_corr_all <- p.adjust(all_results$p.value, method = mc_comp_method)
all_results[] <- lapply(all_results, function(x) if(is.numeric(x)) round(x, 3) else x)

tinytable::tt(all_results) |> 
  tinytable::save_tt(file.path(obj[["home_dir"]], "histories/all_comps.html"), 
                               overwrite = TRUE)
```

(Unweighted for illustrative purposes only)

```{r}
all_results_uw <- rbind(results_t1_uw$comps,
                     results_t2_uw$comps,
                     results_d0_uw$comps,
                     results_d1_uw$comps,
                     results_d2_uw$comps,
                     results_d3_uw$comps,
                     results_d4_uw$comps,
                     results_d5_uw$comps)
all_results_uw$p_corr_all <- p.adjust(all_results_uw$p.value, method = mc_comp_method)
all_results_uw[] <- lapply(all_results_uw, function(x) if(is.numeric(x)) round(x, 3) else x)

tinytable::tt(all_results_uw) |>
  tinytable::save_tt(file.path(obj[["home_dir"]], "/histories/all_comps_uw.html"),
                               overwrite = TRUE)
```

<br> <br> <br>

# Package Citations

We are grateful to the authors of many existing packages that *devMSMs* draws from!

```{r}
grateful::cite_packages(out.dir = obj[["home_dir"]], 
                        omit = c("devMSMs", "devMSMsHelpers"), 
                        out.format = "docx")
```
