---
title: "examplePipeline"
author: "Isa Stallworthy"
date: "8/26/2022"
output: html_document
---

Example pipeline running the MSM functions in sequence 

## User inputs
```{r}
data_path="/Users/isabella/Desktop/BSL Lab/FLP/New/data_long.csv"
home_dir= "/Users/isabella/Desktop/BSL Lab/FLP/New/"

ID="s_id"

time_pts=c(6, 15, 24, 35, 58)
time_var="TAge"
epochs=data.table::data.table( #epochs corresponding to hypotheses of interest
  early=c(6, 15),
  middle=c(24, 35),
  late=58)

missing=-9999

exposures=c("HOMEETA1", "INR", "ESETA1", "CTSETA1")
exposure_time_pts=c(6, 15, 25, 36)

outcomes=c("pcx_sensitive", "pcx_pomd", "pcx_nomd", "pcx_qrel", "pcx_CompTwo")
outcome_time_pts=54

```


### 1. Define your causal question 
What exposures are you interested in in relation to what outcomes?

```{r}

defineCausalQuestion(exposures, exposure_time_pts, outcomes, outcome_time_pts)

```


### 2. Format data & identify potential covariate confounding variables within the data provided
Make sure that your dataset includes all possible available variables that could be potential confounding variables for your causal question of interest. 
```{r}
factor_covariates=c("NC","TcSex2", "TcBlac2") 

assessDataStruct(data_path, ID, time_pts, time_var, missing, exposures, outcomes)

```


### 3. Identify which covariates could potentially confound the relationship between exposure and outcome that will be used to create balancing weights
```{r}
#visually inspect potential covariates to identify those that are:
# - time-varying
# - factors
# - not supposed to be included

print(potential_covariates)

#list out covariates that are time-varying
time_varying_covariates= c("INR", "WIND", "pcx_sensitive", "BSI","ESETA1", "HOMEETA1", "CTSETA1", "pcx_pos", "pcx_neg", "pcx_pomd","pcx_nomd", "pcx_qrel", "GrosPay1", "pcx_engaged", "pcx_CompTwo")

#list out any variables that should be excluded based on practical or theoretical reasons; default set to none
exclude_covariates=c("Index1")


identifyCovariates(home_dir, data, time_pts, potential_covariates, factor_covariates, time_varying_covariates, exclude_covariates)

```


### 4. Impute dataset so that each time point has complete data
5 imputed datasets will be generated and weights will be generated on each imputed dataset (and combined at the end)
```{r}

#list continous variables here; all others assumed ordinal
continuous_variables=c("GrosPay1", "SwghtLB", "RMomAgeU", "RMAge1st", "RWghtLb", "IBQDnovm", "MDI", "INR", "WIND", "ECBQ", "HOMEETA1", "CTSETA1", "ESETA1", "pcx_sensitive", "pcx_pos", "pcx_neg", "pcx_engaged", "pcx_sensitive", "SHCircCm", "pcx_CompTwo", "pcx_pomd", "pcx_nomd", "pcx_qrel", "BSI")

imputeData(home_dir, ID, data_to_impute, continuous_variables)
```


### 5. Format each imputed dataset for calculating balancing weights 
```{r}
#list any time-varying variables that should not be present because of planned missingness design 
time_var_exclude=c("WIND.1", "WIND.24", "BSI.35", "pcx_CompTwo.6", "pcx_CompTwo.1", "pcx_CompTwo.2", "pcx_CompTwo.35")

formatForWeights(ID, home_dir, imputed_datasets, time_varying_covariates, time_pts, just_imputed="yes")
```


### 6. Create formulae for calculating balancing weights for each exposure at each time point 
```{r}
#when creating a balancing formula at each time point, we need to remove potential colliders at that time point. list potential colliders (i.e., variables that could cause both the treatment and the outcome) at each time point below. these will be EXCLUDED from balancing only at the time point of the exposure as balancing on colliders can lead to problematic outcomes. the code automatically removes other outcomes. 
potential_colliders=c()

createForms(exposures, outcomes, time_pts, potential_colliders)

```

### 7. Creating balancing weights for each exposure at each time point, for each imputed dataset
```{r}

createWeights(exposures, time_pts, balance_thresh=0.12, ATT=0, iterations=1000, standardize=FALSE, method="exact", twostep=TRUE, sample.weights=NULL, baseline.forumula=NULL, diff.formula=NULL)

```

### 8. Assess balance and determine variables for which balance could not be achieved for use as covariates in final model
```{r}

assessBalance(home_dir, exposures, time_pts, just_made_weights="no")

```

### 9. Condense weights to create one per person
```{r}

condenseWeights(home_dir, exposures, time_pts)

```

### 10. Truncate weights to avoid heavy tails
```{r}

truncateWeights(exposures, percentile_cutoff=0.90)


```


