---
title: "Recommended Preliminary Steps Prior to using devMSMs with Longitudinal Data"
author: "Isabella C. Stallworthy", "Meriah L. DeJoseph", "Emily R. Padrutt", "Noah Greifer", "Daniel Berry"
date: "`r Sys.Date()`"
output: html_document
---

Please review the accompanying manuscript for a full conceptual and practical introduction to MSMs in the context of developmental data. Please also see the vignettes on the *devMSMs* website for step-by-step guidance on the use of this code: https://istallworthy.github.io/devMSMs/index.html. 

Headings denote accompanying website sections and steps. We suggest using the interactive outline tool (located above the Console) for ease of navigation.

This notebook walks a user through the steps recomennded prior to using *devMSMs*. 

# *Preliminary Steps Vignette*
https://istallworthy.github.io/devMSMs/articles/Preliminary_Steps.html

Functions from devMSMsHelpers: https://github.com/istallworthy/devMSMsHelpers  
  

## Getting Started
You will always need to install the *devMSMs helper* functions from Github (*devMSMsHelpers*).

```{r}
install.packages("devtools", quiet = TRUE)
library(devtools)

install_github("istallworthy/devMSMsHelpers", quiet = TRUE)
library(devMSMsHelpers)
```


## Specify Required Package Core Inputs
The user should change all fields in this code chunk to match their home directory and wide data.

```{r}
# set seed for reproducibility 

set.seed(1234)

# required if you wish to use save.out = TRUE in the functions

home_dir <- NA
home_dir <- '/Users/isabella/Library/CloudStorage/Box-Box/BSL General/MSMs/testing/isa' # note: no / after

# required
exposure <- "ESETA1"

# required
exposure_time_pts <- c(6, 15, 24, 35, 58)

# required
outcome <- "StrDif_Tot.58"

#required; list in wide format
tv_confounders <- c("SAAmylase.6","SAAmylase.15", "SAAmylase.24",
                    "MDI.6", "MDI.15",                                            
                    "RHasSO.6", "RHasSO.15", "RHasSO.24","RHasSO.35", "RHasSO.58",                                         
                    "WndNbrhood.6","WndNbrhood.24", "WndNbrhood.35", "WndNbrhood.58",                                       
                    "IBRAttn.6", "IBRAttn.15", "IBRAttn.24",                                   
                    "B18Raw.6", "B18Raw.15", "B18Raw.24", "B18Raw.58",                                           
                    "HOMEETA1.6", "HOMEETA1.15", "HOMEETA1.24", "HOMEETA1.35", "HOMEETA1.58",                               
                    "InRatioCor.6", "InRatioCor.15", "InRatioCor.24", "InRatioCor.35", "InRatioCor.58",                         
                    "ESETA1.6", "ESETA1.15", "ESETA1.24", "ESETA1.35", "ESETA1.58",  # exposure variables required               
                    "CORTB.6", "CORTB.15", "CORTB.24",                                                                  
                    "EARS_TJo.24", "EARS_TJo.35",                                        
                    "LESMnPos.24", "LESMnPos.35",                                  
                    "LESMnNeg.24", "LESMnNeg.35",       
                    "StrDif_Tot.35", "StrDif_Tot.58",    
                    "fscore.35", "fscore.58"
                    # , "ESETA1.6:B18Raw.6", "ESETA1.6:B18Raw.15:RHasSO.6", "state:EARS_TJo.35" #testing interactions
) 

#required
ti_confounders <- c("state", "BioDadInHH2", "PmAge2", "PmBlac2", "TcBlac2", "PmMrSt2", "PmEd2", "KFASTScr",
                    "RMomAgeU", "RHealth", "HomeOwnd", "SWghtLB", "SurpPreg", "SmokTotl", "DrnkFreq",
                    "peri_health", "caregiv_health", "gov_assist"
                    #, "state:SmokTotl", "PmAge2:PmBlac2", "PmAge2:PmEd2" #testing interaction terms
)
```







### P4. Optional: Specify Exposure Epochs
Users have the option of specifying exposure epochs, or menaingful units of developmental time that encompass the time points at which exposure was measured. We recommend users specify exposure epochs consistently throughout this workflow.  

```{r}
# change this to match your data/theory 

epochs <- data.frame(epochs = c("Infancy", # list user-specified names
                                "Toddlerhood", 
                                "Childhood"), 
                     values = I(list(c(6, 15), # list corresponding exposure time points from data
                                     c(24, 35), 
                                     c(58)
                     ))) 
```



### P5. Recommended: Specify and Inspect Exposure Histories
Regardless of data formatting and exposure epochs, we highly recommend users inspect their data and the distribution of their sample throughout the exposure histories that will be used in subsequent steps.  

#### P5a. Identify High and Low Cutoff Values for Continuous Exposures
For continuously distributed exposures, users have the option to specify cutoffs for demarcating high and low levels of exposure, respectively. We recommend users specify high and low cutoffs consistently throughout this workflow.  

```{r}
# optional list of quantiles specifying high and low cutoff values, respectively, for continuous exposures; default is median

hi_lo_cut <- c(0.6, 0.3) # empirical example 
```


#### P5b. Specify Hypotheses-Relevant Exposure Histories
We recommend users specify the exposure histories that are relevant for testing their hypotheses. We suggest users specify reference and comparison histories consistently throughout this workflow.  

```{r}
# optional reference history (required if comparisons are specified)

reference <- NULL
reference <- "l-l-l" # empirical example final choice
reference <- c("l-l-l", "l-l-h")


# optional comparison history/histories

comparison <- NULL 
comparison <- "h-h-h" # single
comparison <- c("h-h-h", "h-l-l", "l-l-h", "h-h-l", "l-h-h") # empirical example final choice
```



