---
title: "Recommended Workflow for using devMSMs with Longitudinal Data"
author: "Isabella C. Stallworthy", "Meriah L. DeJoseph", "Emily R. Padrutt", "Noah Greifer", "Daniel Berry"
date: "`r Sys.Date()`"
output: html_document
---
This example workflow provides more detail on the application of *devMSMs* to the empirical example in the manuscript, xxxx, which can be found here: 

Please review the accompanying manuscript for a full conceptual and practical introduction to MSMs in the context of developmental data. Please also see the vignettes on the *devMSMs* website for step-by-step guidance on the use of this code: https://istallworthy.github.io/devMSMs/index.html. 

We highly recommend first implementing the *Data Requirements* and *Specify Core Inputs* Vignettes prior to using this workflow. 

Headings denote accompanying website sections and steps. We suggest using the interactive outline tool (located above the Console) for ease of navigation.

The code in each code chunk is set up to showcase all possible inputs to each function (both required and optional) to aid the user's use of the full range of package functionality. Example possible values for the optional input are shown for each function, including a NULL/NA option if the user does not wish to specify an optional input. The user should select one of each optional input values. Alternatively, the user could modify the call to the function and remove the optional input argument(s) entirely. 

Please see the website vignettes and/or type `?functionName` into the console for more guidance on the arguments for each function. These two sources should match but let me know if you see discrepancies. 

Functions from devMSMs: https://github.com/istallworthy/devMSMs


# *Installation*
https://istallworthy.github.io/devMSMs/index.html 

Until *devMSMs* is available on CRAN, you will need to install it directly from Github (https://github.com/istallworthy/devMSMs), as shown below.  

```{r}

install.packages("devtools", quiet = TRUE)
library(devtools)

install_github("istallworthy/devMSMs", quiet = TRUE)
library(devMSMs)

install_github("istallworthy/devMSMsHelpers", quiet = TRUE)
library(devMSMsHelpers)

```


# PRELIMINARY SETUP PHASE: Specify & Inspect Core Package Inputs

## P1. Read in Wide Data
We highly recommend first implementing the *Data Preparation Vignette* https://istallworthy.github.io/devMSMs/articles/Preliminary_Steps.html before assigning to the variable, `data`, one of the following wide data formats (see Figure 1) for use in the package:  

* a single data frame of data in wide format with no missing data  

* a mids object (output from mice::mice()) of data imputed in wide format  

* a list of data imputed in wide format as data frames  

See the Data Preparation vignette for more detail.
  
    
We first read in data that has been multiply imputed 5 times using the *mice* package using the random forest technique. 
```{r}

data <- readRDS("/Users/isabella/Library/CloudStorage/Box-Box/BSL General/MSMs/testing/testing data/continuous outcome/continuous exposure/FLP_wide_imputed_mids.rds") 

```


## P2. Create MSM Object
The first step is to create an initial object by specifying the core variables and data for use with the package. See the Specify Core Inputs vignette for more information. 

Below, we specify data, exposure, time invariatn and time-varying confounders, as well as exposure epochs. 
```{r}

set.seed(1234)

obj <- initMSM(
  data,
  exposure = c("ESETA1.6", "ESETA1.15", "ESETA1.24", "ESETA1.35", "ESETA1.58"), 
  ti_conf =  c("state", "BioDadInHH2", "PmAge2", "PmBlac2", "TcBlac2", "PmMrSt2", "PmEd2", "KFASTScr",
               "RMomAgeU", "RHealth", "HomeOwnd", "SWghtLB", "SurpPreg", "SmokTotl", "DrnkFreq",
               "peri_health", "caregiv_health", "gov_assist"),
  tv_conf = c("SAAmylase.6","SAAmylase.15", "SAAmylase.24", 
              "MDI.6", "MDI.15",                                            
              "RHasSO.6", "RHasSO.15", "RHasSO.24","RHasSO.35", "RHasSO.58",                                         
              "WndNbrhood.6","WndNbrhood.24", "WndNbrhood.35", "WndNbrhood.58",                                       
              "IBRAttn.6", "IBRAttn.15", "IBRAttn.24",                                   
              "B18Raw.6", "B18Raw.15", "B18Raw.24", "B18Raw.58",                                           
              "HOMEETA1.6", "HOMEETA1.15", "HOMEETA1.24", "HOMEETA1.35", "HOMEETA1.58",                               
              "InRatioCor.6", "InRatioCor.15", "InRatioCor.24", "InRatioCor.35", "InRatioCor.58",                         
              "CORTB.6", "CORTB.15", "CORTB.24",                                                                  
              "EARS_TJo.24", "EARS_TJo.35",                                        
              "LESMnPos.24", "LESMnPos.35",                                  
              "LESMnNeg.24", "LESMnNeg.35",       
              "StrDif_Tot.35", 
              "fscore.35", "fscore.58"),
  epoch = c("Infancy", "Infancy", "Toddlerhood", "Toddlerhood", "Childhood"),
  sep = "\\.",
  home_dir = '/Users/isabella/Library/CloudStorage/Box-Box/BSL General/MSMs/testing/isa' 
)

```

Inspect the msm object
```{r}

print(obj)

```


## P3. Required for Continuous Exposures: Identify High and Low Cutoff Values 
See the Specify Core Inputs vignette for more information. 

Below, we specify the 60th and 30th percentiles to demarcate high and low levels of economic strain exposure, respectively.  
```{r}

hi_lo_cut <- c(0.6, 0.3)

```
<br>


## P4. Recommended: Specify Hypotheses-Relevant Exposure Histories 
See the Specify Core Inputs vignette for more information. 

Below, we specify low economic strain at all epochs ("l-l-l") as our reference event in comparison to high levels at all epochs ("h-h-h") as well as all histories that contain 1 dose of exposure to high economic strain at different epochs.  
```{r}

reference <- c("l-l-l")

comparison <- c("h-h-h", 
                "l-l-h", 
                "h-l-l", 
                "l-h-l")
```

<br>

## P5. Recommended: Inspect Exposure Histories and Data

The helper `inspectData()` function outputs the following files into the home directory: a correlation plot of all variables in the dataset, tables of exposure and outcome descriptive statistics, and two summary tables of the confounders considered at each time point.

```{r, eval = FALSE}

outcome <-  "StrDif_Tot.58"

inspectData(data = data, 
            obj = obj, 
            outcome = outcome, 
            reference = reference, 
            comparison = comparison, 
            # home_dir = home_dir, 
            verbose = TRUE,
            save.out = FALSE)
```
Here, we see summaries of the data types as well as reasonable cell counts in each of our specified histories, for each imputed dataset. 

<br>
<br>

# *Workflow: Continuous Exposure Vignette*
https://istallworthy.github.io/devMSMs/articles/Workflow_Continuous_Exposure.html

## PHASE 1: Confounder Adjustment
The first phase of the MSM process is focused on eliminating confounding of the relation between exposure and outcome.  

### STEP 1: Create Full Balancing Formulas & Conduct Pre-Balance Checking
The first step is to create full balancing formulas that reflect all measured confounders at each exposure time point.  

#### 1a. Create Full Balancing Formulas at each Exposure Time Point
Users have the option to specify concurrent confounders to retain and we recommend doing so consistently throughout this workflow. 

Please see the Customize Balancing Formulas Vignette at the link below for more detail on custom formulas. 
https://istallworthy.github.io/devMSMs/articles/Customize_Balancing_Formulas.html  

```{r}

type <- "full"

full_formulas <- createFormulas(obj = obj, 
                                type = type, 
                                save.out = TRUE)

```

Inspect
```{r}

print(full_formulas)

```


#### 1b. Conduct Exploratory Pre-Balance Assessment
The next step is to examine initial imbalance between confounders and exposure prior to IPTW weighting. Users have the option to specify balance threshold(s), which we recommend doing consistently throughout this workflow.  

```{r}

balance_thresh <- c(0.05, 0.1) # empirical example 

imp_conf <- c("InRatioCor.6", "InRatioCor.15", "InRatioCor.24", "InRatioCor.35",
              "InRatioCor.58", "PmEd2") # empirical example

prebalance_stats <- assessBalance(obj = obj, 
                                  data = data,
                                  balance_thresh = balance_thresh, 
                                  imp_conf = imp_conf, 
                                  save.out = TRUE)

```

Inspect
```{r}

summary(prebalance_stats, 
        i = 1, 
        t = 1, 
        save.out = TRUE)

plot(prebalance_stats, 
     i = 1, 
     t = 1, 
     save.out = TRUE)

print(prebalance_stats, 
      i = 2, 
      t = 1, 
      sve.out = TRUE) # NOTE: "F_b" for one ti covar? covariate?
```


### STEP 2: Create Simplified Balancing Formulas & Determine Optimal Weighting Method
The next step is to specify shorter, simplified balancing formula for the purposes of determining the weighting method optimal for the data.   

#### 2a. Create Simplified Balancing Formulas
First, create shorter, simplified balancing formulas at each exposure time point.  

```{r}

type <- "short" 

short_formulas <- createFormulas(obj = obj, 
                                 type = type, 
                                 save.out = TRUE) 

```

Inspect
```{r}

print(short_formulas)

```


#### 2b. Create IPTW Balancing Weights Using Multiple Weighting Methods
We recommend users use the short formulas to create IPTW weights using all the available weighting methods: "glm", "gbm", "bart", "super", and "cbps". 

```{r}

formulas <- short_formulas

method <- "cbps" 

weights.cbps <- createWeights(obj = obj, 
                              data = data, 
                              formulas = formulas, 
                              method = method, 
                              maxit = 1, # for testing purposes only; makes it run faster 
                              save.out = TRUE)
```

Inspect
```{r}

print(weights.cbps, 
      i = 2)

plot(weights.cbps, 
     i = 1, s
     ave.out = TRUE)

summary(weights.cbps[[1]]) 

```

We then fit weights for all other available weighting methods. 
```{r}
method <- "glm"

weights.glm <- createWeights(obj = obj, 
                             data = data, 
                             formulas = formulas, 
                             method = method,
                             save.out = TRUE)
print(weights.glm, 
      i = 2)

plot(weights.glm, 
     i = 2)

method <- "gbm"

weights.gbm <- createWeights(obj = obj, 
                             data = data, 
                             formulas = formulas, 
                             method = method,
                             save.out = TRUE)
print(weights.gbm, 
      i = 2)

plot(weights.gbm, 
     i = 2)

method <- "bart"

weights.bart <- createWeights(obj = obj, 
                              data = data, 
                              formulas = formulas, 
                              method = method,
                              save.out = TRUE)
print(weights.bart, 
      i = 2)

plot(weights.bart, 
     i = 2)

method <- "super"

weights.super <- createWeights(obj = obj, 
                               data = data, 
                               formulas = formulas, 
                               method = method,
                               save.out = TRUE)
print(weights.super, 
      i = 2)

plot(weights.super, 
     i = 2)

```


#### 2c. Assess All Weighting Methods to Determine Optimal Method
The next step is to assess balance for each weighting method and for the user to determine the optimal weighting method.  

```{r}

balance_thresh <- c(0.05, 0.1) # empirical example 

imp_conf <- c("InRatioCor.6", "InRatioCor.15", "InRatioCor.24", 
              "InRatioCor.35", "InRatioCor.58", "PmEd2") # empirical example 

weights <- weights.cbps 

balance_stats.cbps <- assessBalance(data = data, 
                                    obj = obj, 
                                    weights = weights,
                                    imp_conf = imp_conf, 
                                    balance_thresh = balance_thresh,
                                    save.out = TRUE)

# to load in from file (output from saving above!)
# balance_stats.cbps <- readRDS("/Users/isabella/Library/CloudStorage/Box-Box/BSL General/MSMs/testing/isa/balance/weighted-exposure_A-method_cbps.rds")

weights <- weights.glm
balance_stats.glm <- assessBalance(data = data, 
                                   obj = obj, 
                                   weights = weights,
                                   imp_conf = imp_conf, 
                                   balance_thresh = balance_thresh,
                                   save.out = TRUE)

weights <- weights.gbm
balance_stats.gbm <- assessBalance(data = data, 
                                   obj = obj, 
                                   weights = weights,
                                   imp_conf = imp_conf, 
                                   balance_thresh = balance_thresh,
                                   save.out = TRUE)


weights <- weights.bart
balance_stats.bart <- assessBalance(data = data, 
                                    obj = obj, 
                                    weights = weights,
                                    imp_conf = imp_conf, 
                                    balance_thresh = balance_thresh,
                                    save.out = TRUE)


weights <- weights.super 
balance_stats.super <- assessBalance(data = data, 
                                     obj = obj, 
                                     weights = weights,
                                     imp_conf = imp_conf, 
                                     balance_thresh = balance_thresh,
                                     save.out = TRUE)

```

We then compare the balance of the different weighting methods. 

```{r}

summary(balance_stats.cbps)

summary(balance_stats.glm)

summary(balance_stats.gbm)

summary(balance_stats.bart)

summary(balance_stats.super)

```

```{r}

summary(balance_stats.cbps[[1]])

summary(balance_stats.glm[[1]])

summary(balance_stats.gbm[[1]])

summary(balance_stats.bart[[1]])

summary(balance_stats.super[[1]])

```

We identify xx as the weighting method that most successfully reduces imbalance. 



### STEP 3: Create Updated Formulas & Re-Specify Weights Using Optimal Weighting Method
The next step is to update the short balancing formulas with any imbalanced confounders, and re-specify the IPTW weights.  

#### 3a. Inspect Balance of Best-Performing Weighting Method
First, assess how well each of the IPTW achieve balance for all measured confounders. 

```{r}

summary(balance_stats.cbps)

plot(balance_stats.cbps, 
     t = 1)

print(balance_stats.cbps) 

```


#### 3b. Update Simplified Formulas 
Next, update the short formulas with any imbalanced confounders.  

```{r}

type <- "update"

bal_stats <- balance_stats.cbps

updated_formulas <- createFormulas(obj = obj, 
                                   type = type,
                                   bal_stats = bal_stats,
                                   save.out = TRUE)
```

Inspect
```{r}

print(updated_formulas)

```


#### 3c. Create Final Balancing Weights
Next, create final balancing weights using the optimal weighting method and updated balancing formulas.  

```{r}

formulas <- updated_formulas

method <- "cbps" 

final_weights <- createWeights(data = data, 
                               obj = obj, 
                               method = method, 
                               formulas = formulas,
                               maxit = 1, # just for testing purposes 
                               save.out = TRUE)

```

Inspect
```{r}
print(final_weights, 
      i = 1)

summary(final_weights[[1]])

plot(final_weights, 
     i = 1)
```


#### 3d. Trim Final Balancing Weights
Next, trim the final balancing weights to reduce the heavy right tail.  

##### Main
First, trim the main weights.  

```{r}

quantile <- 0.95 # empirical example 

weights <- final_weights

trim_weights <- trimWeights(obj = obj,
                            weights = weights, 
                            at = quantile, 
                            save.out = TRUE)

```

Inspect
```{r}
print(trim_weights, 
      i = 1)

plot(trim_weights, 
     i = 1, 
     save.out = TRUE)

```


##### Sensitvity Analyses
Next, conduct sensitivity analyses using two other quantile values.  

```{r}

quantile <- 0.92 

trim_weights.s1 <- trimWeights(obj = obj,
                               weights = weights, 
                               at = quantile,
                               save.out = TRUE)

quantile <- 0.98

trim_weights.s2 <- trimWeights(obj = obj,
                               weights = weights, 
                               at = quantile,
                               save.out = TRUE)
```



### STEP 4: Conduct Final Balance Assessment
Next, conduct a final balance assessment using all measured confounders (i.e., the full balancing formulas).  

#### Main
First, conduct the main balance assessment.  

```{r}

balance_thresh <- c(0.05, 0.1) # empirical example 

imp_conf <- c("InRatioCor.6", "InRatioCor.15", "InRatioCor.24", 
              "InRatioCor.35", "InRatioCor.58", "PmEd2") # empirical example 

weights <- trim_weights

final_balance_stats <- assessBalance(data = data, 
                                     obj = obj, 
                                     weights = weights,
                                     save.out = TRUE)

```

Inspect
```{r}

summary(final_balance_stats, 
        save.out = TRUE)

plot(final_balance_stats, 
     t = 1, 
     save.out = TRUE)

print(final_balance_stats, 
      save.out = TRUE)


# manually list remaining imbalanced covariates that are time-invariant or time-varying at t=1 for use in Step 5

covariates <- c("PmEd2", "ESETA1.6", "InRatioCor.6")
```


#### Sensitvity Analyses
Next, conduct the recommended specifying sensitivity analyses to match the main analyses above.  

```{r}

weights <- trim_weights.s1

final_balance_stats.s1 <- assessBalance(data = data, 
                                        obj = obj,
                                        weights = weights,
                                        save.out = "weighted-exposure_A-method_cbps_s1.rds")

weights <- trim_weights.s2

final_balance_stats.s2 <- assessBalance(data = data, 
                                        obj = obj, 
                                        weights = weights,
                                        save.out = "weighted-exposure_A-method_cbps_s2.rds")
```



## PHASE 2: Assess Substantive Associations between Exposure and Outcome
Lastly, having attenuated confounder associations, we model substantive associations.  

### STEP 5: Fit Weighted Model & Summarize & Visualize Results

#### 5a. Select and Fit a Weighted Outcome Model
First, select and fit a marginal outcome model.  

##### Main
First, fit the main model.  

```{r}

family <- gaussian

link <- "identity" 

int_order <- 3

weights <- trim_weights

model <- "m1"

outcome <- "StrDif_Tot.58"

models <- fitModel(data = data, 
                   obj = obj, 
                   weights = weights, 
                   outcome = outcome, 
                   model = model,
                   covariates = covariates, 
                   int_order = int_order,
                   save.out = TRUE) 

```

Inspect
```{r}

print(models, 
      i = 1, 
      save.out = TRUE)

```



##### Sensitvity Analyses
Next, fit the recommended  specifying sensitivity analyses to match the main analyses above.   

```{r}
weights <- trim_weights.s1

models.s1 <- fitModel(data = data, 
                      obj = obj, 
                      weights = weights, 
                      outcome = outcome, 
                      model = model,
                      covariates = covariates, 
                      int_order = int_order,
                      save.out = "outcome_D_3-exposure_A-model_m1_s1.rds")

weights <- trim_weights.s2

models.s2 <- fitModel(data = data, 
                      obj = obj, 
                      weights = weights, 
                      outcome = outcome, 
                      model = model,
                      covariates = covariates, 
                      int_order = int_order,
                      save.out = "outcome_D_3-exposure_A-model_m1_s2.rds")

```


#### 5b. Estimate, Compare, & Visualize Model-Predicted Outcome as a Function of History
Lastly, estimate and compare user-specified exposure histories.   

##### Main
First, conduct main comparisons. 

```{r}

hi_lo_cut <- c(0.6, 0.3) # empirical example final choice

reference <- "l-l-l" # empirical example final choice

comparison <- c("h-h-h", "h-l-l", "l-l-h", "h-h-l", "l-h-h") # empirical example final choice

mc_comp_method <- "BH" # empirical example 

dose_level <- "h" # empirical example 

exp_lab <- "Economic Strain" # empirical example 

out_lab <- "Behavior Problems" # empirical example 

colors <- c("Dark2") # empirical example

model <- models 

results <- compareHistories(obj = obj, 
                            fit = model, 
                            comparison = comparison,
                            reference = reference,
                            mc_comp_method = mc_comp_method, 
                            dose = "l",
                            save.out = TRUE)

```

Inspect
```{r}
print(results, 
      save.out = TRUE)

summary(results, 
        save.out = TRUE)

plot(results, 
     save.out = TRUE)
```



##### Sensitvity Analyses
We recommend specifying sensitivity analyses to match the main analyses above. 

```{r}
model <- models.s1 

results.s1 <- compareHistories(obj = obj, 
                               fit = model, 
                               comparison = comparison, 
                               reference = reference, 
                               mc_comp_method = mc_comp_method, 
                               dose = "l",
                               save.out = "outcome_D_3-exposure_A-model_m1_s1.rds")

model <- models.s2 

results.s2 <- compareHistories(obj = obj, 
                               fit = model, 
                               comparison = comparison, 
                               reference = reference,
                               mc_comp_method = mc_comp_method, 
                               dose = "l",
                               save.out = "outcome_D_3-exposure_A-model_m1_s2.rds")

```



# Package Citations
We are grateful to the authors of many existing packages that *devMSMs* draws from!

```{r}
grateful::cite_packages(out.dir = home_dir, 
                        omit = c("devMSMs", "devMSMsHelpers"), 
                        out.format = "docx")
```

