---
title: "Terminology"
date: today
format:
    html:
      df_print: kable
      toc: true
vignette: >
  %\VignetteIndexEntry{Terminology}
  %\VignetteEncoding{UTF-8}s
  %\VignetteEngine{quarto::html}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/"
)

options(repos = c(CRAN = "https://cloud.r-project.org"))

install.packages("markdown")
install.packages("tinytable")

require(tinytable)
require(markdown)
```
<br>

This vignette contains some key definitions for using the <a href="https://istallworthy.github.io/devMSMs/articles/Data_Requirements.html">Data Requirements</a>, <a href="https://istallworthy.github.io/devMSMs/articles/Specify_Core_Inputs.html">Specifying Core Inputs</a>, <a href="https://istallworthy.github.io/devMSMs/articles/Preliminary_Steps.html">Recommended Preliminary Steps</a>, and *Workflows* vignettes. A more thorough introduction to these concepts can be found in the forthcoming manuscript: *[insert link]*.  
<br>  
  
## Key Terminology

```{r}
#| echo: false
#| tbl-cap: "Summary of terminology used in *devMSMs*"
terms <- data.frame(
  Term = c("**Outcome**", "**Exposure**", "**Exposure Time Points**", "**Exposure Epochs**", "**Exposure Histories**", "**Exposure Dosage**", "**Backdoor Path**", "**Confounder**", "**Time-varying confounder**", "**Time invariant confounder**", "**Collider**"),
  Definition = c(
    "Any developmental construct measured at one or more time points including at a final outcome time point",
    "Exposure or experience that constitutes the causal event of interest and is measured at at least two time points, with at least one time point occurring prior to the outcome",
    "Time points in development when the exposure was measured at which balancing formulas will be created",
    "*(optional)* Further delineation of exposure time points into meaningful units of developmental time, each of which could encompass multiple exposure time points, that together constitute exposure main effects in the outcome model and exposure histories",
    "Sequences of relatively high (`'h'`) or low (`'p'`) levels of exposure at each exposure time point or exposure epoch",
    "Total cumulative exposure epochs during which an individual experienced high (or low) levels of exposure across an entire exposure history",
    "Any alternative, non-causal way (other than the direct causal effects from exposure to outcome) that the exposure at a given time point associates with the outcome within a causal diagram, often via confounders",
    "Pre-exposure variable that represents a common cause of exposure at a given time point and outcome; adjusting for all of which successfully blocks all backdoor paths",
    "A confounder measured at one or more time points prior to the exposure time point which reflects a construct that varies in time.",
    "A confounder measured at a single time point prior to the exposure time point which reflects a construct that is stable over time.",
    "A variable that represents a common effect of exposure at a given time point and outcome; adjusting for which opens a backdoor path"
  )
)

tinytable::tt(terms, theme = "striped") |>
  tinytable::format_tt(markdown = TRUE)
```
<br>  

## Balancing Formula

Balancing formula are formula created at each exposure time point, regressing exposure on confounders, for the purpose of creating and/or assessing balancing weights. The general form is given by:
<pre><code>ESETA1.t ~ <span style='color: darkred'>time-invariant</span> confounders + <span style='color: blue'>time-varying</span> confounders + <i>lagged</i> <span style='color: seagreen;'>outcomes</span> + <i>lagged</i> <span style='color: purple;'>outcomes</span></code></pre>

Below is a table that summarizes the three formula `type`'s that will be used.

Table 2. 

```{r}
#| echo: false
#| tbl-cap: "Summary of the formula terminology used in the *devMSMs* package for creating balancing weights and assessing their performance in devMSMs (*Workflows* vignettes Steps 4-6). The different kinds of formulas vary with respect to their inclusion of time-varying confounders and the `type` specification to create them using the `createFormulas()` function."
balance_formula <- data.frame(
  `Term` = c(
    '*Full balancing formula*\n\n`type = "full"`',
    '*Simplified balancing formula*\n\n`type = "short"`',
    '*Updated balancing formula*\n\n`type = "update"`'
  ),
  `Definition` = c(
    "Balancing formula containing all <span style='color: darkred;'>time invariant</span>, <i>all</i> lagged <span style='color: blue;'>time-varying</span> confounders, and lagged <span style='color: seagreen'>exposure</span> values. Used for pre-balance (<i>Workflows</i> vignettes Step 4b) and final balance assessment (<i>Workflows</i> vignettes Step 6).\n\n<pre><code>ESETA1.24 ~ <span style='color: blue;'>B18Raw.15</span> + <span style='color: blue;'>B18Raw.6</span> + <span style='color: darkred;'>BioDadInHH2</span> + <span style='color: blue;'>CORTB.15</span> + <span style='color: blue;'>CORTB.6</span> + <span style='color: darkred;'>DrnkFreq</span> + <span style='color: seagreen;'>ES.15</span> + <span style='color: seagreen;'>ES.6</span> + <span style='color: darkred;'>gov_assist</span> + <span style='color: blue;'>HOMEETA1.15</span> + <span style='color: blue;'>HOMEETA1.6</span> + <span style='color: darkred;'>HomeOwnd</span> ...</code></pre>",
    "Balancing formula containing all <span style='color: darkred;'>time invariant</span>, <i>only t-1</i> lagged <span style='color: blue;'>time-varying</span> confounders, and lagged <span style='color: seagreen'>exposure</span> values. Used for creating and assessing preliminary balancing weights to determine optimal weighting method (<i>Workflows</i> vignettes Step 5b).\n\n<pre><code>ESETA1.24 ~ <span style='color: blue;'>B18Raw.15</span> + <span style='color: darkred;'>BioDadInHH2</span> + <span style='color: blue;'>CORTB.15</span> + <span style='color: darkred;'>DrnkFreq</span> + <span style='color: seagreen;'>ES.15</span> + <span style='color: darkred;'>gov_assist</span> + <span style='color: blue;'>HOMEETA1.15</span> + <span style='color: darkred;'>HomeOwnd</span> ...</code></pre>",
    "Balancing formula containing all <span style='color: darkred;'>time invariant</span>, <i>only t-1</i> lagged <span style='color: blue;'>time-varying</span> confounders, as well as any <i>imbalanced</i> <span style='color: blue;'>time-varying</span> confounders, and lagged <span style='color: seagreen'>exposure</span> levels <i>at t-1+</i> Used to create final balancing weights (<i>Workflows<i> vignettes Step 6).\n\n<pre><code>ESETA1.24 ~ <span style='color: blue;'>B18Raw.15</span> + <span style='color: darkred;'>BioDadInHH2</span> + <span style='color: blue;'>CORTB.15</span> + <span style='color: darkred;'>DrnkFreq</span> + <span style='color: seagreen;'>ES.6</span> + <span style='color: seagreen;'>ES.15</span> + <span style='color: darkred;'>gov_assist</span> + <span style='color: blue;'>HOMEETA1.15</span> + <span style='color: darkred;'>HomeOwnd</span> ...</code></pre>"
  )
)

tinytable::tt(balance_formula) |>
  tinytable::format_tt(markdown = TRUE)
```
<br>  

## 6 core functions of *devMSMs*

There are 6 key functions that are used in the *devMSMs* workflow.

```{r}
#| echo: false
#| tbl-cap: "Summary of the 7 core functions of the *devMSMs* package."
#| results: asis

main_functions <- data.frame(
  Function = c(
    "[`initMSM()`](/reference/initMSM)",
    "[`createFormulas()`](/reference/createFormulas.html)",
    "[`assessBalance()`](/reference/assessBalance)",
    "[`createWeights()`](/reference/createWeights)",
    "[`trimWeights()`](/reference/trimWeights)",
    "[`fitModel()`](/reference/fitModel)",
    "[`compareHistories()`](/reference/compareHistories)"
  ),
  Purpose = c(
    "Creates MSM object with core variables for use in all other functions.",
    "Creates balancing formulas at each exposure time point relating exposure to confounders.",
    "Assesses confounder balance at each exposure time point according to best practices (Jackson, 2016).",
    "Creates balancing weights for each person (wrapper for `weightit::weightitMSM`).",
    "Trims balancing weights to account for heavy tails.",
    "Fits user-selected, weighted marginal model relating exposure to outcome.",
    "Estimate, compare, & visualize model-predicted outcomes as a function of exposure history."
  ),
  `Required Input` = c(
    "`exposure`, `time invariant confounders`, `time-varying confounders`)",
    "`msmObject`, `type` (`'full'`, `'short'`, `'update'`)",
    "`data`, `msmObject`",
    "`data`, `msmObject`",
    "`data`, `msmObject`",
    "`data`, `msmObject`, `weights`, `outcome`, `model` (`m0`-`m3`)",
    "`data`, `msmObject`, `fitted model`"
  ),
  Output = c(
    "msmObject",
    "List of balancing formulas for each exposure time point.",

    "List of balance statistics. Tables & love plots displaying balance statistics at each exposure time point.",
    "Weights object. Histogram & summary of balancing weights.",
    "Trimmed weights object.Histogram & summary of trimmed weight.",
    "Fitted model object. Omnibus test & table of model evidence.",
    "Tables of estimated values and comparisons. Boxplot."

  )
)

tinytable::tt(main_functions, theme = "striped") |>
  tinytable::format_tt(markdown = TRUE)
```


